---
title: "Homework 8"
author: "Mike James"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
H9P1 <- read.csv("./H9P1.csv", sep = ",", header = TRUE)
H9P2 <- read.csv("./H9P2.csv", sep = ",", header = TRUE)
```

## Problem I

> A plant physiologist is interested in how wild chile plants respond to
environmental variability, specifically atmospheric RELATIVE HUMIDITY.  

> Our plant physiologist measured the WATER POTENTIAL at the tip of three leaves for each of 38 plants along a relative
humidity gradient in Arizona.  

> He hypothesized that there is a CUBIC RELATIONSHIP between the
RELATIVE HUMIDITY in the air and the WATER POTENTIAL in chile plants.  

> He averaged the three leaves to minimize measurement error, thus only one reading per plant, and transformed all
the data to meet the assumptions of the test.

***

### Justification

Since the relative humidity is along a controlled gradient, this is a Model I ANOVA utilizing fixed factors.

$H_0:$ No variance between the relative humidty and the water potential in the leaves.  

***

### Results

```{r, echo=FALSE}
# head(H9P1)
summary(H9P1)
t.test(H9P1$asin.relative.humidity, H9P1$H2o.potential)
```  
***  
    
We can immediately see that the means are not equal.

***

```{r, echo=FALSE}
scatterplot(H2O.potential ~ asin.relative.humidity, data=H9P1, xlab = "Relative Humidity", ylab = "Water Potential at the Leaf Tips")
```

***

Our scatter plot shows the data looks to be cubic in nature.

***

```{r, echo=FALSE}
plot(lm(H2O.potential ~ asin.relative.humidity, data=H9P1), which = 1)
```

***

We *might* have a couple of outliers, but they do not appear to affect the outcome.

***

```{r, echo=FALSE}
H9P1.lm1 <- lm(H2O.potential~asin.relative.humidity, H9P1)
H9P1.lm2 <- lm(H2O.potential~asin.relative.humidity + I(asin.relative.humidity^2), H9P1)
H9P1.lm3 <- lm(H2O.potential~asin.relative.humidity + I(asin.relative.humidity^2) + I(asin.relative.humidity^3), H9P1)
H9P1.lm4 <- lm(H2O.potential~asin.relative.humidity + I(asin.relative.humidity^2) + I(asin.relative.humidity^3) + I(asin.relative.humidity^4), H9P1)

anova(H9P1.lm1, H9P1.lm2)
anova(H9P1.lm2, H9P1.lm3)
```

***

Up through this point, all data points to valid variance.

***

```{r, echo=FALSE}
anova(H9P1.lm3, H9P1.lm4)
```

***

Here we can see that fitting the data to the 4th degree is not statistically significant (p = 0.4164).

***

```{r, echo=FALSE}
plot(H9P1.lm3, which=1)
```

***

While there may be a possible outlier in the upper ranges of the relative humidity, they do not change the outcome much, if at all.

***

```{r, echo=FALSE}
attach(H9P1)

plot(H2O.potential~asin.relative.humidity, data = H9P1, pch=16, xlab="Relative Humidity", ylab="Water Potential", main="Relative Humidity Vs. Water Potential in Leaves")
x <- seq(0,10,l=1000)
points(x, predict(H9P1.lm3, data.frame(asin.relative.humidity=x)), type="l", col="red")

detach(H9P1)
```


### Conclusion

According to ANOVA, the data is -- in fact -- cubic (F = 69.748, p <<< 0.05) thus allowing us to reject the null hypothesis.  

As the relative humidity increases, the water potential in the leaves increases until the relative humidity reaches a value of ~5.54, at which point the water potential begins to drop off.

***
***
***

## Problem II

> Our botanist, still interested in chiles' biology, collected seeds from the highest elevation
populations and brought them back to the lab. The seeds were weighed, planted, and
followed for two weeks to determine which ones GERMINATED (1), and which ones DID NOT GERMINATE (0).  

> Our intrepid botanist wanted to predict, based on SEED MASS, which seeds will emerge in two
weeks and which ones wouldn’t.

### Justification

All samples were taken from the highest elevation; thus, this is a fixed factor Model I ANOVA.

$H_0:$ No correlation between the seed mass and whether the seeds germinate. This is a fixed effect in that the seeds either germinated or they did not.

***

### Results

```{r, echo=FALSE}
#head(H9P2)
summary(H9P2)
t.test(H9P2$seed.mass, H9P2$germination)
```

```{r, echo=FALSE}
boxplot(seed.mass ~ as.factor(germination), data=H9P2, xlab = "Germination", ylab="Seed Mass (mg)", main = "Germination Based on Seed Mass")
```

***

The boxplot shows a marked cutoff at 40mg, at which point seeds with a mass **below** 40mg were not likely to germinate, while those with a mass > 40mg were able to germinate.

***

```{r, echo=FALSE}
H9P2.lm <- lm(seed.mass ~ as.factor(germination), data=H9P2)
plot(H9P2.lm, which = 1)
```
***

Here out residuals are nicely separated nicely.

***


```{r, echo=FALSE}
H9P2.glm <- glm(H9P2$germination~H9P2$seed.mass, family=binomial, data=H9P2)
summary(H9P2.glm)
```

***

Here we can see the z-values the coefficients are statistically valid (intercept: p = 0.0188, seed.mass: p = 0.0106)

***


```{r, echo=FALSE}
anova(H9P2.lm, test="Chisq")
```

***

Our $\chi^2$ test shows variance in the seed mass with respect to germination (F = 12.324, p = 0.001534).

***

```{r, echo=FALSE}
H9P2.glm <- glm(germination~seed.mass, family=binomial, data=H9P2)

xs <- seq(0,100,l=30)
H9P2.predict <- predict(H9P2.glm, type="response", se=T, newdata=data.frame(seed.mass=xs))

plot(germination~seed.mass, data=H9P2, xlab="", ylab="", pch=16, axes=F, main = "Does Seed Mass Affect Germination?")
points(H9P2.predict$fit~xs, type="l", col="red")
lines(H9P2.predict$fit+H9P2.predict$se.fit ~ xs, col="gray", type="l", lty=2)
lines(H9P2.predict$fit-H9P2.predict$se.fit ~ xs, col="gray", type="l", lty=2)

mtext(expression("Germinated"), 2, line=3)
axis(2,at=seq(0,1), c("No","Yes"))
mtext("Seed Mass",1, line=3)
axis(1)
box(bty="l")

```

### Conclusion

We are able to reject the null hypothesis as there is obvious variance between seed mass and whether the seeds germinate.