---
title: "LAB_002"
author: "Michael James"
date: "January 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set the working directory to the Dropbox folder

```{r}
setwd("~/Dropbox/SLU/BIOMETRY/LAB/Data")
```

Let's go ahead and load a library.

```{r}
library(ISwR)
```

Import our "crabs.csv" into a data.frame called "crabs" and display the summary info

```{r}
crabs <- read.csv("crabs.csv")
summary(crabs)
```
Now, Camilo added "crabs" with attach(), so I will too

```{r}
attach(crabs)
```

A quick histogram, then an adjusted histogram (to make the empty bin "lump" with the previous) with axis labels and a proper title
```{r}
hist(BODYWT)
hist(BODYWT, breaks=5, xlab = "Crab Body Weight (mg)", ylab = "Number of observed crabs", main = "Observed Crab Body Weight Distribution", col = "red", border = "white")
```

Next up, we're going to look at a scatterplot

```{r}
plot(GILLWT,BODYWT, xlab = "Crab Gill Weight (mg)", ylab = "Crab Weight (mg)", col = "orange", main = "Gill weight VS. Body Weight")
```

Box plot time. By the way:  

> '~' means "AS A FUNCTION OF  

```{r}
boxplot(GILLWT ~ SEX, xlab = "sex", ylab = "Gill Weight (mg)", main = "Sex VS. Gill Weight (mg)")
boxplot(BODYWT ~ SEX, xlab = "sex", ylab = "Body Weight (mg)", main = "Sex VS. Body Weight (mg)", ylim = c(5,30), col = "red")
```

**The dots in the Body Weight VS. Sex are outliers.**

***

# Indexing  

Figure out what type of data we have in crabs, then show specific rows and columns

```{r}
(class(crabs))
(gillWt <- crabs[7:11,2])
(bodyWt <- crabs[7:11,1])
(ALL <- crabs[7:11,1:2])
```

Removing what I just made so I can do the next part

```{r}
rm(gillWt)
rm(bodyWt)
rm(ALL)
(crabs$SEX)
(maleCrabs <- crabs[crabs$SEX == "M", ])
```

Plot Gill Weight VS. Body Weight of males

```{r}
attach(maleCrabs)
plot(GILLWT,BODYWT, xlab = "Crab Gill Weight (mg)", ylab = "Crab Weight (mg)", col = "blue", main = "Gill Weight VS. Body Weight of MALES ONLY")
detach(maleCrabs)
```

Transform Gill Weight with a log function

```{r}
plot(log10(BODYWT), GILLWT, xlab = "Crab Body Weight (mg)", ylab = "Crab Gill (mg)", col = "blue", main = "Sexual Dimorphism")
detach(crabs)
```

Sample T-test

```{r}
attach(crabs)
t.test(log10(GILLWT) ~ SEX)
t.test(log10(BODYWT) ~ SEX)
detach(crabs)
```