---
title: "LAB10_Urchin_Example"
author: "Judith Rodriguez"
date: "April 15, 2019"
output: html_document
---

Last HW
A marine biologist is interested on how the protein level in the diets of sea urchins affects growth (as measured by inter-radial suture). He maintained urchins in diets that were either low or high in protein content, as well as those individuals that occurred in the wild (reference). Are there differences in growth rate among the groups?

Perform full factorial and Ancova in the Dataset
```{r}
urchin_data <- read.csv("h6p1(1).csv",sep=",")
attach(urchin_data)
summary(urchin_data)
```

```{r}
DIET <- factor(DIET, labels = c("High","Low","reference"))
summary(urchin_data)
```

```{r}
plot(IV,SUTW,pch=as.numeric(DIET))
plot(IV, SUTW, pch=as.numeric(DIET), log="xy")
```

```{r}
scatterplot(SUTW~IV|DIET, urchin_data) 
```
```{r}
scatterplot(log(SUTW)~log(IV)|DIET, urchin_data)
```


```{r}
plot(log10(IV),log10(SUTW),pch=as.numeric(DIET))

#Split data by glucose	
tethym.high <- urchin_data[DIET=="High", ]
tethym.low <- urchin_data[DIET=="Low", ] 
tethym.reference <- urchin_data[DIET=="reference", ] 

#Plot separate regression lines	
lm.high <- lm(log10(SUTW)~log10(IV), data=tethym.high)
lm.low <- lm(log10(SUTW)~log10(IV), data=tethym.low)
lm.reference <- lm(log10(SUTW)~log10(IV), data=tethym.reference)

abline(lm.high)
abline(lm.low)
abline(lm.reference)
```

HW - ANCOVA
PERFORM AN ANCOVA TO THE URCHIN DATA.
THE HIGHER DATA IS DIFFERENT FROM EVERYBODY ELSES WE NEED TO PERFORM THE REST PF THE DATA. TAKE IS OUT AND SEE IF THE NEXT IS THE DIFFERENT FROM THE OTHER.

