---
title: "LAB_005"
author: "Mike James"
date: "2/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- save(limpets, file="limpets.RData") -->
<!-- setwd(~/myproject/data/) -->
<!-- load(file=~/myproject/data/"whatever.RData") -->

# One- and Two-Sample Tests

### One-Sample t-test

H~0~: Null Hypothesis
H~a~: Sample mean is different form the theorhetical mean  

> Parametric: 

```{r}
daily.intake <- c(5260,5470,5640,6180,6390,6515,6805,7515,7515,8230,8770)
mean(daily.intake)
sd(daily.intake)
quantile(daily.intake)
t.test(daily.intake, mu=7725)
```

$6754 \pm1142$

Is **our** mean indistinguishable from that of Lipitor®? 

$$t = \frac{\bar{x} - \mu_{0}{SEM}}$$

$$P(Tt\le t) = \int_{-\inf}^{t}f(t)$$

Two-sided is the $\pm$SD at the upper and lower boundaries.

```{r}
daily.intake <- c(5260,5470,5640,6180,6390,6515,6805,7515,7515,8230,8770)
mean(daily.intake)
sd(daily.intake)
quantile(daily.intake)
t.test(daily.intake, mu=7725, alternative="greater")
```

Here, the t-test is wildly different because we set up the "alternative hypothesis."

***

# Wilcoxon Signed-Rank Test  

If your data is small, do *not* worry about transforming it.  

```{r}
daily.intake <- c(5260,5470,5640,6180,6390,6515,6805,7515,7515,8230,8770)
mean(daily.intake)
sd(daily.intake)
quantile(daily.intake)
wilcox.test(daily.intake, mu=7725)
wilcox.test(daily.intake, mu=7725, alternative="greater")
```

This isn't as powerful as the parametric tests

### Two-Sample t-test

$$t = \frac{\bar{x_{2}} - \bar{x_{1}}{SEDM}$$

$$SEDM = \sqrt{SEM_{1}^{2} - SEM_{2}^{2}}$$

```{r}
library(ISwR)
(attach(energy))
t.test(expend~stature)
t.test(expend~stature, var.equal=T)
```

And, of course, we do this with the non-parametric Wilcoxon Test

```{r}
wilcox.test(expend~stature)
```

### Paired t-test

```{r}
(attach(intake))
t.test(pre, post, paired=T)
wilcox.test(pre, post, paired=T)
```

### Testing Assumptions

Is the data **normally** distributed?

H~0~: Yes they are
H~a~: No, they are not

```{r}
shapiro.test(daily.intake)
tapply(expend, stature, shapiro.test)
```

Are the variances even?

```{r}
var.test(expend~stature)
```

$Pop \rightarrow Sample$
$\bar{x} \leadsto \mu$
$s \leadsto \sigma$  

***

```{r}
library(readr)
(LeafData <- read_delim("~/Dropbox/SLU/BIOMETRY/LAB/Data/LeafData.tsv","\t", escape_double = FALSE, trim_ws = TRUE))
summary(LeafData)

# Make sure to attach the data so you can reference the headers
attach(LeafData)
t.test(HEIGHT~PARK)

# Determine if the data is normal
var.test(log10(HEIGHT)~PARK)
tapply(log10(HEIGHT), PARK, shapiro.test)

# Boxplot
boxplot(HEIGHT~PARK)
boxplot(log10(HEIGHT)~PARK)

# Wilcoxon
wilcox.test(HEIGHT~PARK)
```

The differences are in variances, not the means.  

In Hawn, more short plants than tall. Similar to the "Bill Gates statistically makes us millionaires when he walks in the room."  

***

# Next weeks homework

1. LEAF.NUMBER~PARK
2. LEAF.W~LIFE.STAGE

## Three Sections
### ONE: Justification
* Which test?
* Why?
    * Normality
    * Homogeneity of VARs
    * Graph (boxplot())  
    
### TWO: Actual Test and Results
* t-tests
* graphs (could be the same, could be the log10 transformed version)

### THREE: Conclusion
* Accept/Reject because of...???


