---
title: "LAB_010_ggplot"
author: "Mike James"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(Rmisc)
```

***

### ggPlot2

Here's the data given to us for the lab (ggplot.example.data.csv):

```{r}
# ggplot.example.data <- read.csv()
head(ggplot.example.data)
```

The plants are inoculated with something and then specific treatments are used.  

The population can be split into known quantities; thus, they are a fixed effect.  

Genet v. Ramit: Genet is a plant that comes from the roots of a plant / Ramit is an individual plant.  

***

#### Example

We have an orchard with four farms. We have fruit flies. 

We treat two of sections of trees in the orchard...

```{r}
(data.SE <- summarySE(ggplot.example.data, measure = "Log.Seed.set", groupvars = c("Population", "Infection.state")))
```

Now, to plot it with ggplot2:

```{r}
ggplot(data.SE, aes(x = Population, y = Log.Seed.set, group = Infection.state, color = Infection.state)) +
         geom_errorbar(aes(ymin = Log.Seed.set - se, ymax = Log.Seed.set + se), color = "black", width = 0.1) +
         geom_point() +
         geom_line() + # IF there is a correlation
         ylab("Log(Seed Set) ± se") + # if you need the xlab, do the same thing
         scale_color_hue(name = "Infection State") + # This gives a custom title to the legend
         theme(axis.title.x = element_text(size = rel(1)), 
               axis.title.y = element_text(size = rel(1)), 
               axis.text = element_text(size = rel(1.0)),
               legend.text = element_text(size = rel(1)),
               legend.title = element_text(size = rel(1)),
               legend.justification = c(5.7,-3.5),
               legend.position = c(1,0))
```

### Something about homework

Generate the appropriate graph using stage **and** treatment with all of the elements.

***

# THIS IS DUE WEDNESDAY!!!!!!!

