---
title: "LAB_003"
author: "Michael James"
date: "February 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

# R packages

* This is from ISwR  

```{r}
sample(1:40,5)
sample(1:10,5)
sample(1:5,5)
```

*These were sampled **without** replacement*

```{r}
sample(1:5,5,replace=T)
```

*These were sample **WITH** replacement*  

***

# Combinatorics

$(_k^n) = \frac{n!}{k!(n-k)!}$  

```{r}
# 3 black, 2 reds
# As of this point, this is incorrect. Should be something with choose(52,x)
black <- choose(52,3)
red <- choose(52,2)
(total <- black + red)
```

# Density

Flipping a coin 10 times and getting 3 heads
Flipping a coin 10 times and getting 7 heads

```{r}
dbinom(3, size=10, prob=0.5)
dbinom(7, size=10, prob=0.5)  
```

> Post-hoc Probability: "Let's make a deal"  

```{r}
x <- 0:10
plot(x,dbinom(x, size=10, prob=0.5), type="h")
```


> Normal p.d.f.  

```{r}
x <- seq(-4,4,0.1)
plot(x, dnorm(x), type="l")
curve(dnorm(x), from = -4, to = 4)
```

# Cumulative Distribution Function 

**See lecture notes**

# Quantiles

```{r}
# Proportion of individuals, mean we're looking at, standard deviation
qnorm(0.95, 132, 13)
```

Example of a 95% Confidence Interval

$\bar{X}\pm95$%


```{r}
xbar <- 83
sigma <- 12
n <- 5
(sem <-sigma/sqrt(n))
xbar + sem + qnorm(0.0250)
xbar + sem + qnorm(0.975)
```

***

# Random numbers

```{r}
(rnorm(10))
(rnorm(10, mean=7, sd=5))
(rbinom(10, size = 20, prob = 0.5))
```

