---
title: "LAST LAB"
author: "Mike James"
date: "5/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Information Reduction

```{r}
head(AquaticInsectReduced)
summary(AquaticInsectReduced)
```

    Rivers - Month
    
> SPx: from(x, 100)  


    PCA ~ Reduction Analysis  
    
**Never stop with PCA** as it has *lots* of faults.

**When importing**: make sure you use "Row Names > Use First Column"  

> Number of Principal Componenets = Number of Objects (???)

> PCs ~ axes [hypercube]

Allow PCs that explain **at least** 5%

```{r}
aquatics.pca1 <- prcomp(AquaticInsectReduced, scale=TRUE)
biplot(aquatics.pca1)
summary(aquatics.pca1)
aquatics.pca1
```

89.741% of the variation is explain in the first 3 samples, so we can dump the other 14 (?) samples.

Scale this data from weighted to unweighted


```{r}
library(vegan)
aquatics.rda <- rda(AquaticInsect[,2:71], scale = TRUE)
biplot(aquatics.rda)
```


```{r}
data(mtcars)
head(mtcars)
```

```{r}
data(mtcars)
# head(mtcars)
# summary(mtcars)
pca.cars <- prcomp(mtcars, scale=TRUE)
biplot(pca.cars)
summary(pca.cars)
```

```{r}
# RDA: Redundancy Analysis

rda.cars <- rda(mtcars, scale=TRUE)
biplot(rda.cars)

# 2 lines close to one another "pretty much" the same data. Keep the one with the greatest number in the list
```


***

# metaMDS: Nonmetric Multidimensional Scaling with Stable Solution from Random Starts, Axis Scaling, and SPecies Scores

### This will replace (?) PCA

```{r}
insects.nmds <- metaMDS(AquaticInsectReduced)
plot(insects.nmds, type="t")
```

PCA = regression
MDS = correlation

MDS can give you a "very naive null hypothesis."

MDS is just for Visualization; not scientific.



