---
title: "Homework_006"
author: "Mike James"
date: "4/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISwR)
library(RColorBrewer)
library(ggplot2)
data <- read.csv("h6p1(1).csv", sep=",")
head(data)
summary(data)
setwd("~/Dropbox/SLU/BIOMETRY/LAB/Homework/006")
```
***
***

*A marine biologist is interested on how the protein level **(DIET)** in the diets of sea urchins affects growth (as measured by inter-radial suture**(SUTW)**). He maintained urchins in diets that were either low or high in protein content, as well as those individuals that occurred in the wild (reference).* 

***Are there differences in growth rate among the groups?***

***
***
***

```{r}
dataHigh <- data[data$DIET == "High", ]
dataLow <- data[data$DIET == "Low", ]
```

```{r}
ggplot(dataLow, mapping = aes(IV, SUTW)) +
  ggtitle("Plate IV Size VS. Intra-radial Suture Size for a Low Protein Diet") +
  geom_point() +
  xlab("Plate IV Size") +
  ylab("Intra-radial Suture Size") +
  scale_color_hue(name = "Diet") +
  geom_smooth(method='lm', formula = y~x)
```

```{r}
ggplot(dataHigh, mapping = aes(IV, SUTW)) +
  ggtitle("Plate IV Size VS. Intra-radial Suture Size for a High Protein Diet") +
  geom_point() +
  xlab("Plate IV Size") +
  ylab("Intra-radial Suture Size") +
  scale_color_hue(name = "Diet") +
  geom_smooth(method='lm', formula = y~x)
```

Both regressions appear to have similar slopes, but I'm not so sure the graphs are linear.


***
***
***

```{r}
ggplot(dataLow, mapping = aes(IV, SUTW)) +
  ggtitle("Plate IV Size VS. Intra-radial Suture Size for a Low Protein Diet") +
  geom_point() +
  xlab("Plate IV Size") +
  ylab("Intra-radial Suture Size") +
  scale_color_hue(name = "Diet") +
  geom_smooth(method='lm', formula = y~poly(x,2,raw=TRUE))
```
```{r}
ggplot(dataHigh, mapping = aes(IV, SUTW)) +
  ggtitle("Plate IV Size VS. Intra-radial Suture Size for a High Protein Diet") +
  geom_point() +
  xlab("Plate IV Size") +
  ylab("Intra-radial Suture Size") +
  scale_color_hue(name = "Diet") +
  geom_smooth(method='lm', formula = y~poly(x,2,raw=TRUE))
```

```{r}
poly_data.lm <- lm(data$IV ~ poly(data$SUTW, 2))
residuals <- residuals(poly_data.lm)
```

### Predictions

Now is the part of the experiment where we're done with our math, we **think** we have some idea how this works, and we should be able to predict what will happen based on this data. 

```{r warning = FALSE}
predict(poly_data.lm, int ="c") # Confidence intervals
predict(poly_data.lm, int="p") # Prediction intervals
```
<!--

[//]: # ```{r warning=FALSE}
x <- data.frame(6)
please <- data.frame(poly_data.lm$IV)
predict.glm(please, newdata=x, int="p")
[//]: # ```


***
***

#```{r}
#options(na.action=na.exclude)
#pred.frame <- data.frame(poly_data.lm$SUTW)
#pp <- predict(data=data, int="p", newdata=pred.frame)
#pc <- predict(data=data, int="c", newdata=pred.frame)
#pred.SUTW <- pred.frame$SUTW
#plot(SUTW, IV, ylim=range(IV, pp, na.rm = TRUE))
#matlines(pred.SUTW, pc, lty = c(3,1,1), col=c("purple","green","blue"))
#matlines(pred.SUTW, pp, lty = 1, col="red")
```

-->

# Results

No, they do not have different growth rates based on diet.

P.S. I ran out of time, hence the shit quality.